Article data analysis

### 2018-10-04 notebooks/article_notebook.Rmd
minor changes

```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable)
```

# Import data
```{r Import data}
isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

@@

# ARIBA results analysis
```{r ARIBA analysis}
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"
ac_genes <- c("qnr","aac","oqx")
mut_genes <- c("gyr","par","marR","soxS","tolC","acrR")

@@

```{r}
per_species <- isolate_data %>%
  group_by(species) %>%
  count()

@@

>>>>

# Dependencies
This notebook is using the following packages:

- dplyr

- ggplot2

- purrr
```{r load libraries, message=FALSE, warning=FALSE}
# Libraries
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, ggplot2, tidyr, cowplot, viridis, forcats, radiant.data, purrr, phangorn, gridExtra, ggtree, htmlTable)
```

```{r, include=FALSE}
id <- c("2017-01-3941-1-S", "2016-17-98-1-S", "2016-17-610-1-S", "2016-17-604-1-S", 
"2016-17-565-1-S", "2016-17-550-1-S", "2016-17-548-1-S", "2016-17-546-1-S", 
"2016-17-485-1-S", "2016-17-469-1-S", "2016-17-423-1-S", "2016-17-412-1-S", 
"2016-17-408-1-T", "2016-17-408-1-S", "2016-17-393-1-S", "2016-17-391-1-S", 
"2016-17-388-1-T", "2016-17-388-1-S", "2016-17-377-1-S", "2016-17-363-1-S", 
"2016-17-352-1-S", "2016-17-343-1-S", "2016-17-340-1-S", "2016-17-336-1-S", 
"2016-17-329-1-S", "2016-17-299-1-S", "2016-17-298-1-S", "2016-17-294-1-S", 
"2016-17-292-1-S", "2016-17-288-1-S", "2016-17-281-1-T", "2016-17-281-1-S", 
"2016-17-275-1-S", "2016-17-271-1-S", "2016-17-27-1-S", "2016-17-264-1-S", 
"2016-17-263-1-S", "2016-17-257-1-S", "2016-17-256-1-S", "2016-17-242-1-S", 
"2016-17-225-1-S", "2016-17-224-1-S", "2016-17-189-1-T", "2016-17-183-1-S", 
"2016-17-174-1-T", "2016-17-174-1-S", "2016-17-164-1-S", "2016-17-16-1-S", 
"2016-17-148-1-S", "2016-17-124-1-S", "2016-17-114-1-S", "2016-17-105-1-T", 
"2016-17-1-1-S", "2016-02-785-2-S", "2016-02-783-2-T", "2016-02-771-2-S", 
"2016-02-770-2-S", "2016-02-741-2-S", "2016-02-739-2-S", "2016-02-732-2-S", 
"2016-02-727-2-S", "2016-02-726-2-S", "2016-02-725-2-S", "2016-02-723-2-S", 
"2016-02-697-2-S", "2016-02-693-2-T", "2016-02-693-2-S", "2016-02-692-2-S", 
"2016-02-690-2-T", "2016-02-690-2-S", "2016-02-620-2-S", "2016-02-563-2-S", 
"2016-02-546-2-S", "2016-02-543-2-S", "2016-02-531-2-S", "2016-02-522-2-S", 
"2016-02-515-2-T", "2016-02-515-2-S", "2016-02-514-2-S", "2016-02-513-2-T", 
"2016-02-513-2-S", "2016-02-505-2-S", "2016-02-492-2-S", "2016-02-490-2-S", 
"2016-02-486-2-S", "2016-02-447-2-S", "2016-02-446-2-S", "2016-02-428-2-S", 
"2016-02-425-2-S", "2016-02-419-2-S", "2016-02-418-2-S", "2016-02-415-2-S", 
"2016-02-413-2-T", "2016-02-413-2-S", "2016-02-411-2-S", "2016-02-410-2-S", 
"2016-02-409-2-S", "2016-02-408-2-S", "2016-02-402-2-S", "2016-02-381-2-S", 
"2016-02-369-2-S", "2016-02-366-2-S", "2016-02-363-2-S", "2016-02-359-2-S", 
"2016-02-358-2-S", "2016-02-356-2-S", "2016-02-353-2-S", "2016-02-352-2-S", 
"2016-02-351-2-S", "2016-02-348-2-S", "2016-02-345-2-T", "2016-02-345-2-S", 
"2016-02-343-2-S", "2016-02-336-2-S", "2016-02-330-2-S", "2016-02-329-2-S", 
"2016-02-324-2-S", "2016-02-316-2-S", "2016-02-302-2-S", "2016-02-293-2-S", 
"2016-02-281-2-S", "2016-02-278-2-S", "2016-01-3763-1-T", "2016-01-3474-1-T", 
"2015-01-753-1-S", "2015-01-752-1-S", "2015-01-751-1-S", "2015-01-750-1-S", 
"2015-01-749-1-S", "2015-01-5263-1-S", "2015-01-5262-1-S", "2015-01-5177-1-S", 
"2015-01-5141-1-S", "2015-01-5140-1-S", "2015-01-5138-1-S", "2015-01-5022-1-S", 
"2015-01-5020-1-S", "2015-01-5018-1-S", "2015-01-4988-1-S", "2015-01-4934-1-S", 
"2015-01-4932-1-S", "2015-01-4931-1-S", "2015-01-4794-1-S", "2015-01-4788-1-S", 
"2015-01-4742-1-S", "2015-01-467-1-S", "2015-01-466-1-S", "2015-01-463-1-S", 
"2015-01-4559-1-S", "2015-01-4558-1-S", "2015-01-4465-1-S", "2015-01-4463-1-S", 
"2015-01-4462-1-S", "2015-01-4421-1-S", "2015-01-4249-1-S", "2015-01-4037-1-S", 
"2015-01-3915-1-S", "2015-01-3914-1-S", "2015-01-3911-1-S", "2015-01-3887-1-S", 
"2015-01-3640-1-S", "2015-01-3639-1-S", "2015-01-3638-1-S", "2015-01-3548-1-S", 
"2015-01-3545-1-S", "2015-01-3318-1-S", "2015-01-3317-1-S", "2015-01-3102-1-S", 
"2015-01-3004-1-S", "2015-01-3001-1-S", "2015-01-2879-1-S", "2015-01-2878-1-S", 
"2015-01-2839-1-S", "2015-01-2838-1-S", "2015-01-2837-1-S", "2015-01-2836-1-S", 
"2015-01-2835-1-S", "2015-01-2771-1-S", "2015-01-2430-1-S", "2015-01-2429-1-S", 
"2015-01-2426-1-S", "2015-01-2404-1-S", "2015-01-2403-1-S", "2015-01-2309-1-S", 
"2015-01-2166-1-S", "2015-01-2097-1-S", "2015-01-2094-1-S", "2015-01-1955-1-S", 
"2015-01-1954-1-S", "2015-01-1952-1-T", "2015-01-1951-1-S", "2015-01-1950-1-S", 
"2015-01-1949-1-S", "2015-01-1826-1-S", "2015-01-1823-1-S", "2015-01-1661-1-S", 
"2015-01-1566-1-S", "2015-01-1343-1-S", "2014-01-861-1-S", "2014-01-860-1-S", 
"2014-01-856-1-S", "2014-01-7584-1-S", "2014-01-7543-1-S", "2014-01-7512-1-S", 
"2014-01-7490-1-S", "2014-01-7479-1-S", "2014-01-7443-1-S", "2014-01-7442-1-S", 
"2014-01-7394-1-S", "2014-01-7375-1-S", "2014-01-7234-1-S", "2014-01-7207-1-S", 
"2014-01-7136-1-S", "2014-01-7133-1-S", "2014-01-7050-1-S", "2014-01-6926-1-S", 
"2014-01-6924-1-S", "2014-01-6710-1-S", "2014-01-6625-1-S", "2014-01-6530-1-S", 
"2014-01-6291-1-S", "2014-01-6123-1-S", "2014-01-6122-1-S", "2014-01-6044-1-S", 
"2014-01-6043-1-S", "2014-01-6041-1-S", "2014-01-6040-1-S", "2014-01-5917-1-T", 
"2014-01-5917-1-S", "2014-01-5916-1-S", "2014-01-5915-1-S", "2014-01-5912-1-S", 
"2014-01-5792-1-S", "2014-01-5565-1-S", "2014-01-4922-1-T", "2014-01-4921-1-S", 
"2014-01-4651-1-S", "2014-01-4539-1-S", "2014-01-4271-1-S", "2014-01-4270-1-S", 
"2014-01-3802-1-S", "2014-01-3677-1-S", "2014-01-3676-1-S", "2014-01-3212-1-S", 
"2014-01-2996-1-S", "2014-01-2995-1-S", "2014-01-2992-1-S", "2014-01-2773-1-S", 
"2014-01-2728-1-S", "2014-01-2516-1-S", "2014-01-2366-1-S", "2014-01-2362-1-T", 
"2014-01-2145-1-S", "2014-01-2144-1-S", "2014-01-2071-1-S", "2014-01-2070-1-S", 
"2014-01-2069-1-S", "2014-01-1742-1-S", "2014-01-1741-1-S", "2014-01-1649-1-S", 
"2014-01-1268-1-T", "2014-01-1182-1-S", "2013-01-2024-1-T", "2012-01-838-6-T", 
"2011-01-9386-3-T", "2011-01-6927-5-T", "2011-01-3460-5-T", "2011-01-1544-2-T", 
"2010-01-991-7-T", "2009-01-551-1-T", "2009-01-4672-5-T", "2009-01-4618-2-T", 
"2009-01-4415-4-T", "2009-01-413-1-T", "2009-01-4064-1-T", "2009-01-3815-1-T", 
"2009-01-3064-2-T", "2009-01-2696-4-T", "2009-01-2521-3-T", "2009-01-175-4-T", 
"2009-01-175-2-T", "2009-01-1690-1-T", "2007-01-361-1-T", "2006-01-2821-1-T", 
"2006-01-2184-1-T")

```{r Import data, include = FALSE}
isolate_data <- read.table("../data/isolate_data.txt",
                           sep = "\t",
                           header = T,
                           stringsAsFactors = F)

@@

```{r ARIBA analysis, include = FALSE}
megares_report_loc <- "../data/ariba/megares_results"
resfinder_report_loc <- "../data/ariba/resfinder_results"
ac_genes <- c("qnr","aac","oqx")
mut_genes <- c("gyr","par","marR","soxS","tolC","acrR")

@@

## Isolates per species
```{r}
per_species <- isolate_data %>%
  group_by(species) %>%
  count()

@@

# Figures
## Percent QREC isolates with mutations in intrinsic genes
```{r}
gene_names <- names(mut_report)

gene_names <- gene_names[-1]

mut_report %>%
    mutate(
      ref = gsub("(.*?)\\_.+", "\\1", ref),
      ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
      ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)
    ) %>%
    arrange(desc(ref)) %>%
    mutate(id = id) %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, gene_names) %>%
    group_by(key, value, species) %>%
    count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(1)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(1)
    ) +
    scale_x_discrete(limits = c(
      "gyrA", "gyrB", "parC", "parE",
      "marR", "soxS", "tolC"
    )) +
    scale_fill_brewer(palette = "Paired") +
    labs(y = "Percent (%) of isolates") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(
      axis.text.x = element_text(
        angle = 90,
        size = 12,
        hjust = 1,
        vjust = 0.4
      ),
      axis.title.x = element_blank(),
      axis.text.y = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      strip.text = element_text(size = 12)
    ) +
    facet_wrap(~ species)
```

## Percent QREC isolates with Qnr genes per species
```{r}
qnr_genes <- c("qnrA1","qnrB19","qnrB6","qnrB60","qnrS1","qnrS2","qnrS4")

acquired_report %>%
    mutate(
      ref = gsub("(.*?)\\_.+", "\\1", ref),
      ref = sub("^\\d*-?(\\d{4}-.*)", "\\1", ref),
      ref = sub("^(\\d{4}-\\d{2}-\\d*)-1", "\\1", ref)
    ) %>%
    arrange(desc(ref)) %>%
    mutate(id = id) %>%
    left_join(isolate_data, by = "id") %>%
    gather(key, value, qnr_genes) %>%
    group_by(key, value, species) %>%
    count() %>%
    ungroup() %>%
    mutate(value = if_else(value == 1, "Present", "Absent")) %>%
    spread(value, n, fill = 0) %>%
    rowwise() %>%
    mutate(
      Total = Present + Absent,
      Percent = round(Present / Total * 100, 1),
      lwr = round(get_binCI(Present, Total)[1], 1),
      upr = round(get_binCI(Present, Total)[2], 1)
    ) %>%
    ggplot(aes(key, Percent, fill = species)) +
    geom_col(color = "black",
             position = position_dodge(0.9)) +
    geom_errorbar(
      aes(ymin = lwr, ymax = upr),
      width = 0.6,
      position = position_dodge(0.9)
    ) +
    scale_fill_brewer(palette = "Paired") +
    labs(y = "Percent (%) of isolates",
         fill = NULL) +
    theme_classic() +
    theme(
      axis.text = element_text(size = 12),
      axis.title.y = element_text(size = 14),
      legend.text = element_text(size = 12),
      axis.title.x = element_blank(),
      legend.justification = c(0, 1),
      legend.position = c(0.82, 0.97)
    )
```
